|

### 2024-10-28 output/seminar-slideshow.qmd
Include normal probability function in IFA slide

## Análisis factorial de ítems

:::::: columns
:::: column
<br>

$$
  i_k^* = \tau_k + \mathbf{\lambda}_k^T \mathbf{F} + U_k
$$

@@

::: {.fragment fragment-index="1"}
$$
  I_k = \left\{ \begin{matrix}
      1 &si &i_k^* \geq 0\\
      0 &si &i_k^* <    0
    \end{matrix} \right.
$$
:::
::::

::: {.column .fragment fragment-index="1"}
```{r ifa-curve}
#| fig-asp: 1.4
#| cache: false

PROB_AXIS      <- seq(-3, 3, by = .01)
ERROR_POINTS   <- seq(-3, 3, length.out = 5)
ITEM_LOADING   <-  0.35
ITEM_INTERCEPT <- -0.5

@@

# Plot: Upper Panel - Linear regression with Gaussian error distributions
latent_response_plot <- latent_response |>
  ggplot(aes(x = response, y = point, fill = ..x.. > 0, group = point)) +
  geom_density_ridges_gradient(
    scale          =  .2,
    rel_min_height = 0.001,
    color          = PALETTE_BLUE["normal"]
  )                                                                     +
  labs(x = "i~k~^*^", y = NULL)                                         +
  scale_y_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    labels       = NULL,
    limits       = range(ERROR_POINTS) * 1.12,
    expand       = expansion()
    )           +
  scale_x_continuous(breaks = 0)                                        +
  scale_fill_manual(
    values = c(
      `FALSE` = unname(PALETTE_BLUE["white"]),
      `TRUE`  = unname(PALETTE_BLUE["normal"])
    )
  ) +
  theme(
    axis.ticks.x = element_blank(),
    axis.title.y = element_markdown(angle = 90),
    legend.position = "none",
    panel.grid.major = element_line(linetype = "17")
  )                                                                     +
  coord_flip()

# Plot: Lower Panel - Probit curve
probit_plot <- probit_curve_data |>
  ggplot(mapping = aes(x = score, y = prob))                        +
  geom_line(color = PALETTE_BLUE["normal"], linewidth = LINE_WIDTH) +
  labs(x = "E(*i~k~^*^*)", y = "P(I~k~ = 1)")                       +
  scale_y_continuous(
    breaks       = 0:1,
    minor_breaks = NULL,
    limits       = 0:1,
    expand       = expansion(mult = c(0, .05))
  )  +
  scale_x_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    limits = range(ERROR_POINTS) * 1.12,
    expand = expansion()
  ) +
  theme(
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(angle = 90),
    panel.grid.major = element_line(linetype = "17"),
  )

## TODO: Correct error break points and fig height
latent_response_plot / probit_plot
```
:::
::::::

>>>>

## Análisis factorial de ítems {.smaller}

:::::: columns
:::: {.column width="40%"}
$$
  i_k^* = \tau_k + \mathbf{\lambda}_k^T \mathbf{F} + U_k
$$

@@

::: {.fragment fragment-index="1"}
$$
  I_k = \left\{ \begin{matrix}
      1 &si &i_k^* \geq 0\\
      0 &si &i_k^* <    0
    \end{matrix} \right.
$$
:::

::: {.fragment fragment-index="2"}
$$
  \begin{align}
    P(I_k = 1) &= \int_{-\infty}^0 \mathcal{N}(i_k^*) \delta i_k^* \\
               &= \Phi_N (i_k^*)
  \end{align}
$$
:::
::::

:::: {.column width="60%"}
::: {.fragment fragment-index="1"}
```{r ifa-density}
#| fig-asp:   0.5
#| fig-width: 8

PROB_AXIS      <- seq(-3, 3, by = .01)
ERROR_POINTS   <- seq(-3, 3, length.out = 5)
ITEM_LOADING   <-  0.45
ITEM_INTERCEPT <- -0.3

@@

# Plot: Upper Panel - Linear regression with Gaussian error distributions
latent_response |>
  ggplot(
    aes(x = response, y = point, fill = after_stat(x) > 0, group = point)
  )                            +
  geom_density_ridges_gradient(
    scale          =  .2,
    rel_min_height = 0.001,
    color          = PALETTE_BLUE["normal"]
  )                              +
  labs(x = "i~k~^*^", y = NULL)  +
  scale_y_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    labels       = NULL,
    limits       = range(ERROR_POINTS) * 1.12,
    expand       = expansion()
  )           +
  scale_x_continuous(breaks = 0) +
  scale_fill_manual(
    values = c(
      `FALSE` = unname(PALETTE_BLUE["white"]),
      `TRUE`  = unname(PALETTE_BLUE["normal"])
    )
  )                              +
  theme(
    axis.ticks.x = element_blank(),
    axis.title.y = element_markdown(angle = 90),
    legend.position = "none",
    panel.grid.major = element_line(linetype = "17")
  )                              +
  coord_flip()
```
:::

::: {.fragment fragment-index="2"}
```{r ifa-curve}
#| fig-asp:   0.5
#| fig-width: 8

# Plot: Lower Panel - Probit curve
probit_curve_data |>
  ggplot(mapping = aes(x = score, y = prob))                        +
  geom_line(color = PALETTE_BLUE["normal"], linewidth = LINE_WIDTH) +
  labs(x = "E(*i~k~^*^*)", y = "P(I~k~ = 1)")                       +
  scale_y_continuous(
    breaks       = 0:1,
    minor_breaks = NULL,
    limits       = 0:1,
    expand       = expansion(mult = c(0, .05))
  )  +
  scale_x_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    limits = range(ERROR_POINTS) * 1.12,
    expand = expansion()
  ) +
  theme(
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(angle = 90),
    panel.grid.major = element_line(linetype = "17"),
  )
```
:::
::::
::::::
